### 2021-06-23 results.tex
logistic regressions

In this section we detail the findings of our study. We present the results of the first and
second studies in Section~\ref{sec:res-fs} and Section~\ref{sec:res-ss}, respectively. In Section~\ref{sec:discussion} we summarize the
implications of our study.

@@

Altogether, ignoring  \joke, our study reveals that from $58.33$\% (Humanoid)
to $76.04$\% (DroidBot) of the malicious apps investigated in our study can be
detected using the sandboxes generated after running the test case tools with the support of the
DroidFax static analysis algorithms. We also investigate if the use
of the DroidFax static analysis component is statistically significant to identify
malwares in our experiment. To this end, we build a logistic regression model
in the form $Malware \sim Tool + StaticAnalysis + Repetition$. Table~\ref{tab:log-reg1}
shows the results of the logistic regression analysis, highlighting the
Humanoid has a negative and significant impact on malware identification and
the use of DroidFax static analysis has a positive and significant
impact on malware identification.

\begin{table}[ht]
  \centering
     \caption{Results of the Logistic Regression} 
       \begin{small}
     \begin{tabular}{lrlc}
   \toprule
  & \emph{Estimate} & \emph{p-value} & \emph{C.I.} \\ 
   \midrule
   Tool [DroidBot] & 0.1034 & 0.4718 & (-0.133$, $0.340) \\ 
   Tool [DroidMate] & -0.0561 & 0.6955 & (-0.292$, $0.180) \\ 
   Tool [Humanoid] & -0.8910  & 0.0000 $^{***}$ & (-1.131$, $-0.651) \\ 
   Tool [Monkey] & -0.0110 & 0.9390 &  (-0.247$, $0.225)\\ 
   DroidFax static analysis & 0.8867 & 0.0000 $^{***}$ & (0.743$, $1.031)\\ 
   Repetition & -0.0171 & 0.7487 & (-0.105$, $0.071) \\ \midrule
   AIC            & 3001.07     \\
   Num. obs.      & 2304        \\
   \bottomrule
   \multicolumn{4}{l}{\scriptsize{$^{***}$ \emph{p-value} $<0.001$}}
     \end{tabular}
        \end{small}

    \label{tab:log-reg1} 
\end{table}

Besides that, in the first execution (WS), none of the
resulting sandboxes could detect 18 malwares in our dataset ($18.75$\%). According to
the Euphony tool~\cite{hurier2017euphony}, $12$ of these $18$ malwares are \emph{adwares}, $3$ are \emph{trojans}, $2$ are
PUPs (\emph{Potentially Unwanted Program}), and one is an \emph{exploit}.
At this point, an additional question arises: what are the characteristics
of the malwares that have (not) been identified using the mining sandbox approach?
To explore this question, we take advantage of the
\texttt{dex2jar} tool to reverse-engineer all 96 malwares considered in our analysis and
computed the \emph{diffs} of the benign/malicious versions of the APPs.
The results of this activity are available in our replication
package.\footnote{https://github.com/droidxp/paper-replication-package/blob/master/diff/} 
In what follows we dissect a few examples of malwares
that at least one dynamic analysis tool was able to
identify using the mining sandbox approach. After that,
we present the characteristics of a malware that none of the tools
was able to detect. Our goal here is to provide
a lower-level intuition about the classes of malware the
mining sandbox approach is able to detect. A reader that
is not interested in these details could skip to
Section~\ref{sec:res-ss}.

To start with, consider the malicious version of the app \texttt{com.andoop.flyracing}---which both DroidBot and
Humanoid sandboxes could detect in our analysis. 
In this particular case, the malicious version changes the Android Manifest file,
adding permissions to receive and send SMS messages
(Listing~\ref{lst:androidManifest}). With these permissions, malicious apps may get money
fraudulently by sending messages without user confirmation, for instance.

@@

The malicious version of the app \texttt{com.happymaau.MathRef} also changes
the manifest file to require additional permissions as well as change
the behavior of the app (with malicious code). All sandboxes were able to
detect this malware.
In this case, the malicious version of the app changes the Android Manifest file,
requiring permissions to access the network and WiFi states (Listing~\ref{lst:androidManifest2}).
These changes allow an app
to view the status of all networks and make changes to configured WiFi networks.

@@

This pattern of changing the Android Manifest file and including
new method calls characterizes the classes of malwares for which
the mining sandbox approach excels. 
In a different vein, the malicious version of the app
\texttt{com.andoop.flyracing} is among the apps that none of the sandboxes could
detect. Indeed, the malicious version only changes the Android Manifest file,
modifying the meta-data \texttt{ADMOB\_PUBLISHER\_ID}. The AdMob is a monetizing
service provided by Google, and changing the AdMob \emph{publisher identifier} account redirects
the advertisement's revenue to another destination. Based on
this observation, we envision integrating a different approach that reasons
about modifications to the Android manifest file and that might complement the mining sandbox
approach into the task for detecting malwares; since the mining
sandbox approach is not able to detect malicious packages that
do not introduce new method calls for sensitive APIs.

@@

In this second study we used a taint analysis approach to mine differences between the
benign and malicious versions of the 96 Android apps in our dataset. To this end we leverage the FlowDroid
%\kn{ Have we cited flowdroid before. If yes, ignore my comment} 
tool, which tracks how sensitive information flows through the apps using taint analysis algorithms.
Regarding accuracy, the taint analysis approach detected $58$ out of the $96$ pairs in our dataset ($60,42$\%). That is,
using the taint analysis implementation of FlowDroid alone outperforms the Monkey, DroidMate,
and Humanoid sandboxes computed in the second execution (without the DroidFax static analysis
algorithms). This result shows that static analysis algorithms are promising to complement mining
sandboxes, demystifying previous claims in the literature.

@@

Additionally, we investigate if we could benefit from combining the
static analysis strategies from FlowDroid and DroidFax. Figure~\ref{fig:venn-plot2} shows a
Venn-diagram summarizing the results. So, when combining
the results from FlowDroid and DroidFax, we were able to detect
$67$ of the malicious apps ($69.79$\%), a result compatible
to the performance we found as response to the first execution of the
test case generation tools---which also considers the DroidFax
static analysis algorithms. More interesting, from those $67$
malicious apps identified, $33$ pairs had been found by
both tools (FlowDroid and DroidFax), even though they follow
a completely different static analysis approach. Furthermore,
FlowDroid shows to be more effective than DroidFax alone, detecting $25$ malicious
apps that had not been detected by DroidFax (while DroidFax detected $9$
malicious apps that had not been detected by FlowDroid).

>>>>

In this section we detail the findings of our study. We present the results of the first and
second studies in Section~\ref{sec:res-fs} and Section~\ref{sec:res-ss}, respectively. In Section~\ref{sec:implications} we summarize the
implications of our study.

@@

Altogether, ignoring  \joke, our study reveals that from $58.33$\% (Humanoid)
to $76.04$\% (DroidBot) of the malicious apps investigated in our study can be
detected using the sandboxes generated after running the test case tools with the support of the
DroidFax static analysis algorithms. {\color{red}We also investigate if the use
  of the DroidFax static analysis component leads to statistically significant benefit
  on malware identification. To this end, we build a logistic regression model
in the form $Malware \sim Tool + StaticAnalysis + Repetition$. Table~\ref{tab:log-reg1}
shows the results of the logistic regression analysis, highlighting that
(a) Humanoid has a negative, though significant impact on malware identification; and
(b) the use of DroidFax static analysis has a positive and significant
impact on malware identification.}

\begin{table}[ht]
  \centering
     \caption{Results of the Logistic Regression (first study)} 
       \begin{small}
     \begin{tabular}{lrlc}
   \toprule
  & \emph{Estimate} & \emph{p-value} & \emph{C.I.} \\ 
   \midrule
   Tool [DroidBot] & 0.1034 & 0.4718 & (-0.133$, $0.340) \\ 
   Tool [DroidMate] & -0.0561 & 0.6955 & (-0.292$, $0.180) \\ 
   Tool [Humanoid] & -0.8910  & 0.0000 $^{***}$ & (-1.131$, $-0.651) \\ 
   Tool [Monkey] & -0.0110 & 0.9390 &  (-0.247$, $0.225)\\ 
   DroidFax static analysis & 0.8867 & 0.0000 $^{***}$ & (0.743$, $1.031)\\ 
   Repetition & -0.0171 & 0.7487 & (-0.105$, $0.071) \\ \midrule
   AIC            & 3001.07     \\
   Num. obs.      & 2304        \\
   \bottomrule
   \multicolumn{4}{l}{\scriptsize{$^{***}$ \emph{p-value} $<0.001$}}
     \end{tabular}
        \end{small}
    \label{tab:log-reg1} 
\end{table}

Besides that, in the first execution (WS), none of the
resulting sandboxes could detect 18 malwares in our dataset ($18.75$\%). According to
the Euphony tool~\cite{hurier2017euphony}, $12$ of these $18$ malwares are \emph{adwares}, $3$ are \emph{trojans}, $2$ are
PUPs (\emph{Potentially Unwanted Program}), and one is an \emph{exploit}.
At this point, an additional question arises: what are the characteristics
of the malwares that have (not) been identified using the mining sandbox approach?
To explore this question, we take advantage of the
\texttt{dex2jar} tool to reverse-engineer all 96 malwares considered in our analysis and
computed the \emph{diffs} of the benign/malicious versions of the APPs.
The results of this activity are available in our replication
package.\footnote{https://github.com/droidxp/paper-replication-package/blob/master/diff/} 
In what follows we dissect a few examples of malwares
that at least one of the resulting  was able to
identify. After that,
we present the characteristics of a malware that none of the sandboxes
was able to detect. Our goal here is to provide
a lower-level intuition about the classes of malware the
mining sandbox approach is able to detect. A reader that
is not interested in these details could skip to
Section~\ref{sec:res-ss}.

To start with, consider the malicious version of the app \texttt{com.andoop.flyracing}---which both DroidBot and
Humanoid sandboxes could detect in our analysis. 
In this particular case, the malicious version changes the Android Manifest file,
adding permissions to receive and send SMS messages
(Listing~\ref{lst:androidManifest}). Adding these permissions, a malicious app may get money
fraudulently by sending messages without user confirmation, for instance.
The pair \texttt{L:M} indicates
a code segment that appears in line \texttt{L} of the malign (\texttt{M})
version of an app.

@@

The malicious version of the app \texttt{com.happymaau.MathRef} also changes
the Manifest file to require additional permissions as well as change
the behavior of the app (with malicious code). All sandboxes were able to
detect this malware.
In this case, the malicious version of the app changes the Android Manifest file,
requiring permissions to access the network and WiFi states (Listing~\ref{lst:androidManifest2}).
These changes allow an app
to view the status of all networks and make changes to configured WiFi networks.

@@

This pattern of changing the Android Manifest file and including
new method calls characterizes the classes of malwares for which
the mining sandbox approach excels. 
In a different vein, the malicious version of the app
\texttt{com.andoop.flyracing} is among the apps that none of the sandboxes could
detect. Indeed, the malicious version only changes the Android Manifest file,
modifying the meta-data \texttt{ADMOB\_PUBLISHER\_ID}. The AdMob is a monetizing
service provided by Google, and changing the AdMob \emph{publisher identifier} account redirects
the advertisement's revenue to another destination. Based on
this observation, we envision integrating a different approach that reasons
about modifications to the Android Manifest file and that might complement the mining sandbox
approach into the task for detecting malwares; since the mining
sandbox approach is not able to detect malicious packages that
do not introduce new method calls for sensitive APIs.

@@

In this second study we used a taint analysis approach to mine differences between the
benign and malicious versions of the 96 Android apps in our dataset. To this end we leverage the FlowDroid
%\kn{ Have we cited flowdroid before. If yes, ignore my comment} 
tool, which tracks how sensitive information flows through the apps using taint analysis algorithms.
Regarding accuracy, the taint analysis approach detected $58$ out of the $96$ pairs in our dataset ($60,42$\%). That is,
using the taint analysis implementation of FlowDroid alone outperforms the Monkey, DroidMate,
and Humanoid sandboxes computed in the second execution (without the DroidFax static analysis
algorithms). This result shows that static analysis algorithms are promising to complement
the mining sandbox approach.

@@

Additionally, we investigate if we could benefit from combining the
static analysis strategies from FlowDroid and DroidFax. Figure~\ref{fig:venn-plot2} shows a
Venn-diagram summarizing the results. So, when combining
the results from FlowDroid and DroidFax, we were able to detect
$67$ of the malicious apps ($69.79$\%), a result compatible
to the performance we found as response to the first execution of the
test case generation tools---which also considers the DroidFax
static analysis algorithms. More interesting, from those $67$
malicious apps identified, $33$ malwares had been found by
both FlowDroid and DroidFax, even though they follow
a completely different static analysis approach. Furthermore,
FlowDroid shows to be more effective than DroidFax alone, detecting $25$ malicious
apps that had not been detected by DroidFax (while DroidFax detected $9$
malicious apps that had not been detected by FlowDroid). {\color{red}The results
of a logistic regression analysis, considering the
model $Malware \sim Tool$, where Malware is a response
variable indicating if the malware has been detected or not and
Tool is either FlowDroid or the sandbox DroidFax
static analysis component generates, reveals the existence
of a significant difference between
the performance of both tools (see Table~\ref{tab:log-reg2}).}

\begin{table}[ht]
  \centering
     \caption{Results of the Logistic Regression (second study)} 
       \begin{small}
     \begin{tabular}{lrlc}
   \toprule
  & \emph{Estimate} & \emph{p-value} & \emph{C.I.} \\ 
   \midrule
   Tool [FlowDroid] & 0.4229 & 0.0428 $^{**}$ & (0.080$, $0.766) \\
   Tool [DroidFax static analysis component] & -1.2730 & 0.0000 $^{***}$ & (-1.560$, $-0.986) \\ \midrule
   AIC            & 334.61     \\
   Num. obs.      & 288        \\
   \bottomrule
   \multicolumn{4}{l}{\scriptsize{$^{***}$ \emph{p-value} $<0.001$; $^{***}$ \emph{p-value} $<0.05$}}
     \end{tabular}
        \end{small}

    \label{tab:log-reg2} 
\end{table}

### 2021-06-23 setup.tex
logistic regressions

To sum up, in order to analyse the
performance of the test case generation tools (including Joker),
we just have to compare the calls to sensitive APIs made by
the benign and malign versions of the apps, during
the execution of the tools. In the end, we generate
a set of observations, where each observation
contains the tool name, the number of the repetition (in the
range [1..3]), a boolean value reporting the use of the
DroidFax static analysis component, and a boolean value indicating
whether or not the malware has been identified. We use descriptive statistics
and plots to compare the performance of the tools and
answer RQ1 and RQ2. We also use \emph{Logistic Regression} to understand
the relevance of the features (tool, repetition, DroidFax static analysis
component) to predict the malware identification.

@@

We use two metrics in this second study:
the total number of malicious apps FlowDroid is able to
find and the execution time for running the taint analysis algorithm
for each app. Similarly to the first empirical study,
we use descriptive statistics
and plots to compare the performance of the taint analysis and
mining sandbox approaches.

>>>>

To sum up, in order to analyse the
performance of the test case generation tools (including Joker),
we just have to compare the calls to sensitive APIs made by
the benign and malign versions of the apps, during
the execution of the tools. In the end, we generate
a set of observations, where each observation
contains the tool name, the number of the repetition (in the
range [1..3]), a boolean value reporting the use of the
DroidFax static analysis component, and a boolean value indicating
whether or not the malware has been identified. We use descriptive statistics
and plots to compare the performance of the tools and
answer RQ1 and RQ2. We also use \emph{Logistic Regression}~\cite[Chapter~4]{statistical-learning}
to understand the statistical relevance and
the contribution of each feature (tool, repetition, DroidFax static analysis
component) to malware identification.

@@

We use two metrics in this second study:
the total number of malicious apps FlowDroid is able to
find and the execution time for running the taint analysis algorithm
for each app. Similarly to the first empirical study,
we use descriptive statistics
and plots to compare the performance of the taint analysis and
mining sandbox approaches. We also use \emph{Logistic Regression}~\cite[Chapter~4]{statistical-learning}
to better understand the statistical
significance of the benefits of using FlowDroid
(in comparison to the DroidFax static analysis
component only).
